---
title: "Additional calculations for manuscript"
output: html_notebook
---

```{r}
source("/home/rstudio/bicklab-aps/edit/Single_Cell_CHIP/scripts/source_singlet_data.R")
```
```{r}
# Number of cells in TET2 patient
tet2_patient_singlet_data@meta.data %>% nrow()

# Number of mutated and wildtype cells
tet2_patient_singlet_data@meta.data %>%
  select(Clone) %>%
  table()

# Number of mutated and wildtype cells by cell type
tet2_patient_singlet_data@meta.data %>%
  select(Clone, celltype.de) %>%
  table()

# Percentage mutated
4425 / 8585

```
```{r}
# Number of cells in DNMT3A patient
dnmt3a_patient_singlet_data@meta.data %>% nrow()

# Number of mutated and wildtype cells
dnmt3a_patient_singlet_data@meta.data %>%
  select(Clone) %>%
  table()

# Number of mutated and wildtype cells by cell type
dnmt3a_patient_singlet_data@meta.data %>%
  select(Clone, celltype.de) %>%
  table()

# Percentage mutated
872 / 4457
```


```{r}
# Number of differentially expressed genes
setwd("/home/rstudio/bicklab-aps/edit/Single_Cell_CHIP/tables/supplemental/")

# TET2 CD14 monocytes
tet2_data = read_csv("TET2_CD14_Mono_unstim_wilcox.csv")
tet2_data %>%
  filter(abs(avg_log2FC) > 0.25 & p_val_adj < 0.05) %>%
  nrow()

# DNMT3A CD14 monocytes
dnmt3a_data = read_csv("DNMT3A_CD14_Mono_unstim_wilcox.csv")
dnmt3a_data %>%
  filter(abs(avg_log2FC) > 0.25 & p_val_adj < 0.05) %>%
  nrow()

# TET2 CD8 T cells
tet2_data = read_csv("TET2_CD8_T_cell_unstim_wilcox.csv")
tet2_data %>%
  filter(abs(avg_log2FC) > 0.25 & p_val_adj < 0.05) %>%
  nrow()

# DNMT3A CD8 T cells
dnmt3a_data = read_csv("DNMT3A_CD8_T_cell_unstim_wilcox.csv")
dnmt3a_data %>%
  filter(abs(avg_log2FC) > 0.25 & p_val_adj < 0.05) %>%
  nrow()
```
```{r}
# Number of cells in differential expression comparison
singlet_data@meta.data %>% 
  filter(celltype.de == "CD14 Mono" & MUTATION.GROUP == "TET2") %>%
  nrow()

singlet_data@meta.data %>% 
  filter(celltype.de == "CD14 Mono" & MUTATION.GROUP == "DNMT3A") %>%
  nrow()

singlet_data@meta.data %>% 
  filter(celltype.de == "CD14 Mono" & MUTATION.GROUP == "Control") %>%
  nrow()

```

```{r}
# Number of differentially expressed genes - individual patient
setwd("/home/rstudio/bicklab-aps/edit/Single_Cell_CHIP/tables/supplemental/")

# TET2 CD14 monocytes
tet2_data = read_csv("patient_TET2_CD14_Mono_wilcox_unstim.csv")
tet2_data %>%
  filter(abs(avg_log2FC) > 0.25 & p_val_adj < 0.05) %>%
  nrow()

# DNMT3A CD14 monocytes
dnmt3a_data = read_csv("patient_DNMT3A_CD14_Mono_wilcox_unstim.csv")
dnmt3a_data %>%
  filter(abs(avg_log2FC) > 0.25 & p_val_adj < 0.05) %>%
  nrow()

```

```{r}
# Number of aggregate control cells
singlet_data@meta.data %>% 
  filter(MUTATION.GROUP == "Control") %>%
  nrow()
```

```{r}
# Chi-squared test to compare mutant cell counts to expected proportions (based on aggregated control proportions)
setwd("/home/rstudio/bicklab-aps/edit/Single_Cell_CHIP/tables/figure_3/")

summary_stats = singlet_data@meta.data %>%
  filter(MUTATION.GROUP == "Control") %>%
  group_by(celltype.de) %>%
  summarize(count = n()) %>%
  mutate(exp_proportion = count / sum(count)) %>%
  select(celltype.de, exp_proportion)
  
# Number of mutated and wildtype cells by cell type
tet2_counts = tet2_patient_singlet_data@meta.data %>%
  filter(Clone == "Mutant") %>%
  select(celltype.de) %>%
  table() %>%
  as.data.frame()

colnames(tet2_counts) = c("celltype.de", "count")

tet2_comparison_table = full_join(tet2_counts, summary_stats)
tet2_comparison_table[is.na(tet2_comparison_table)] = 0

tet2_chisq_result = chisq.test(tet2_comparison_table$count, p = tet2_comparison_table$exp_proportion)
print(chisq_result)


# Number of mutated and wildtype cells by cell type
dnmt3a_counts = dnmt3a_patient_singlet_data@meta.data %>%
  filter(Clone == "Mutant") %>%
  select(celltype.de) %>%
  table() %>%
  as.data.frame()

colnames(dnmt3a_counts) = c("celltype.de", "count")

dnmt3a_comparison_table = full_join(dnmt3a_counts, summary_stats)
dnmt3a_comparison_table[is.na(dnmt3a_comparison_table)] = 0

dnmt3a_chisq_result = chisq.test(dnmt3a_comparison_table$count, p = dnmt3a_comparison_table$exp_proportion)
print(dnmt3a_chisq_result)
```
```{r}
# T-test to compare lab values for white blood cell counts, absolute monocytes, and percent monocytes between CHIP patients and controls
matched_mutations = singlet_data@meta.data %>%
  select(orig.ident, MUTATION.GROUP) %>%
  unique()

clinical_data = read.xlsx("../input_data/clinical data for brett v3.xlsx") %>%
  inner_join(matched_mutations, by = c("patient_id" = "orig.ident")) %>%
  select(MUTATION.GROUP, wbc, abs_monocytes, `monocytes_%`)

# White blood cells
tet2_values = clinical_data %>% filter(MUTATION.GROUP == "TET2") %>% pull(wbc)
dnmt3a_values = clinical_data %>% filter(MUTATION.GROUP == "DNMT3A") %>% pull(wbc)
control_values = clinical_data %>% filter(MUTATION.GROUP == "Control") %>% pull(wbc)
t.test(tet2_values, control_values)
t.test(dnmt3a_values, control_values)

# Absolute monocytes
tet2_values = clinical_data %>% filter(MUTATION.GROUP == "TET2") %>% pull(`monocytes_%`)
dnmt3a_values = clinical_data %>% filter(MUTATION.GROUP == "DNMT3A") %>% pull(`monocytes_%`)
control_values = clinical_data %>% filter(MUTATION.GROUP == "Control") %>% pull(`monocytes_%`)
t.test(tet2_values, control_values)
t.test(dnmt3a_values, control_values)

# Percent monocytes
tet2_values = clinical_data %>% filter(MUTATION.GROUP == "TET2") %>% pull(abs_monocytes)
dnmt3a_values = clinical_data %>% filter(MUTATION.GROUP == "DNMT3A") %>% pull(abs_monocytes)
control_values = clinical_data %>% filter(MUTATION.GROUP == "Control") %>% pull(abs_monocytes)
t.test(tet2_values, control_values)
t.test(dnmt3a_values, control_values)
```
```{r}
# Regression test to correlate VAF with myeloid skew


# need VAF and myeloid proportion for each patient
mutation_data = read_tsv("../tables/supplemental/mutation_data.tsv") %>%
  select(orig.ident, VAF)

vaf_and_myeloid = singlet_data@meta.data %>%
  inner_join(mutation_data) %>%
  group_by(orig.ident, VAF) %>%
  summarize(cd14_mono_prop = sum(celltype.de == "CD14 Mono") / n())

linear_model = lm(vaf_and_myeloid$VAF ~ vaf_and_myeloid$cd14_mono_prop)
summary(linear_model)

write_tsv(vaf_and_myeloid, "../tables/supplemental/vaf_and_myeloid.tsv")

```


Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
