---
title: "Additional calculations for manuscript"
output: html_notebook
---

```{r}
source("/home/rstudio/bicklab-aps/edit/Single_Cell_CHIP/scripts/source_singlet_data.R")
```
```{r}
# Number of cells in TET2 patient
tet2_patient_singlet_data@meta.data %>% nrow()

# Number of mutated and wildtype cells
tet2_patient_singlet_data@meta.data %>%
  select(Clone) %>%
  table()

# Number of mutated and wildtype cells by cell type
tet2_patient_singlet_data@meta.data %>%
  select(Clone, celltype.de) %>%
  table()

# Percentage mutated
4425 / 8585

```
```{r}
# Number of cells in DNMT3A patient
dnmt3a_patient_singlet_data@meta.data %>% nrow()

# Number of mutated and wildtype cells
dnmt3a_patient_singlet_data@meta.data %>%
  select(Clone) %>%
  table()

# Number of mutated and wildtype cells by cell type
dnmt3a_patient_singlet_data@meta.data %>%
  select(Clone, celltype.de) %>%
  table()

# Percentage mutated
872 / 4457
```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
